---
title: "Assignment 1"
author: "Group 11"
date: "`r Sys.Date()`"
output: html_document
---

# Part B: Synthetic Data
# 1. Introduction

In this part, we generate synthetic data for the proposed study.
The study investigates whether adding a human demonstration to a text
instruction affects the time an embodied AI robot needs to complete a
manipulation task.

The **independent variable** is the instruction condition:
- Text-only
- Text + demonstration

The **dependent variable** is task completion time, measured in seconds.

We generate two datasets:

1. A dataset in which a true effect exists.
2. A dataset in which the null hypothesis is true.

The effect of the human demonstration is controlled by an adjustable
parameter called `effect_size`.

# 2. Load packages

```{r setup, message=FALSE, warning=FALSE}
# Load the packages needed for data manipulation and visualisation
library(tidyverse)
```

# 3. Set seed and Define simulation parameters
```{r}
# Set a seed to for reproducibility
set.seed(42)

# Total number of robot trials
N <- 100

# Number of trials in each experimental condition
n_per_group <- N / 2

# Mean completion time for the text-only condition, in seconds
mean_text <- 30

# Standard deviation of completion time within each condition, in seconds
sd_completion <- 5

# True effect of adding a human demonstration
# A negative value means that the demonstration is expected to
# reduce completion time
# For example, -3 means that the demonstration reduces the
# population mean completion time by 3 seconds
effect_size <- -3

# Mean completion time in the demonstration condition
mean_demo <- mean_text + effect_size
```

# 4.1 Simulate the dataset with a true effect
```{r}
# Generate completion times for robots receiving the text-only instruction
# rnorm() draws random observations from a normal distribution.
# The distribution has a mean of 30 seconds and an SD of 5 seconds
text_effect <- rnorm(
  n = n_per_group,
  mean = mean_text,
  sd = sd_completion
)

# Generate completion times for robots receiving both the text instruction
# and a human demonstration
# The mean is lower by the specified effect size (-3 seconds)
demo_alternative <- rnorm(
  n = n_per_group,
  mean = mean_demo,
  sd = sd_completion
)

# Combine the two groups into one data frame
data_alternative <- tibble(
  condition = c(
    rep("Text-only", n_per_group),
    rep("Text + demonstration", n_per_group)
  ),
  completion_time = c(
    text_effect,
    demo_alternative
  )
)

# Completion time cannot be negative in reality
# We therefore replace any negative simulated values with zero
data_alternative <- data_alternative %>%
  mutate(
    completion_time = pmax(completion_time, 0)
  )

# Display the first few observations
head(data_alternative)
```

# 4.2 Inspect the simulated effect dataset
```{r}
# Calculate descriptive statistics for each experimental condition
data_alternative %>%
  group_by(condition) %>%
  summarise(
    n = n(),
    mean = mean(completion_time),
    sd = sd(completion_time),
    min = min(completion_time),
    max = max(completion_time)
  )
```

# 5.1 Simulate the dataset under null hypothesis
```{r}
# For the null dataset, we set the true effect to zero, which means that 
# the population mean completion time is identical in the two conditions
null_effect <- 0

# Calculate the mean completion time for the demonstration condition
# under the null hypothesis
mean_demo_null <- mean_text + null_effect

# Generate completion times for the text-only condition
text_null <- rnorm(
  n = n_per_group,
  mean = mean_text,
  sd = sd_completion
)

# Generate completion times for the demonstration condition
# Since null_effect = 0, both groups have the same population mean
demo_null <- rnorm(
  n = n_per_group,
  mean = mean_demo_null,
  sd = sd_completion
)

# Combine the two groups into one data frame
data_null <- tibble(
  condition = c(
    rep("Text-only", n_per_group),
    rep("Text + demonstration", n_per_group)
  ),
  completion_time = c(
    text_null,
    demo_null
  )
)

# Ensure that completion times cannot be negative
data_null <- data_null %>%
  mutate(
    completion_time = pmax(completion_time, 0)
  )

# Display the first few observations
head(data_null)
```

# 5.2 Inspect the null dataset
```{r}
# Calculate descriptive statistics for each condition:
# Even though the population means are identical, the sample means
# will usually differ slightly because of random sampling variation
data_null %>%
  group_by(condition) %>%
  summarise(
    n = n(),
    mean = mean(completion_time),
    sd = sd(completion_time),
    min = min(completion_time),
    max = max(completion_time)
  )
```
# 6. Visualize the dataset with a true effect
```{r}
# Plot the distribution of completion times in the two conditions:

# Individual observations are shown as points
# the boxplots summarise the distribution within each condition.
ggplot(
  data_alternative,
  aes(
    x = condition,
    y = completion_time
  )
) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.6) +
  labs(
    title = "Simulated Completion Time: True Effect",
    x = "Instruction condition",
    y = "Completion time (seconds)"
  ) +
  theme_minimal() +
  theme(title = element_text(size = 20)) +
  theme(axis.text.x = element_text(size = 10)) +
  theme(axis.title = element_text(size = 15))
```

# 7. Visualize the null dataset
```{r}
# Visualise the completion times under the null hypothesis:
# The two groups should overlap substantially because their population
# means are identical.
ggplot(
  data_null,
  aes(
    x = condition,
    y = completion_time
  )
) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.6) +
  labs(
    title = "Simulated Completion Time: Null Effect",
    x = "Instruction condition",
    y = "Completion time (seconds)"
  ) +
  theme_minimal() +
  theme(title = element_text(size = 20)) +
  theme(axis.text.x = element_text(size = 10)) +
  theme(axis.title = element_text(size = 15))
```

# 8. Compare the two conditions statistically
```{r}
# Perform the planned independent-samples two-sided t-test
# on the dataset containing a true effect:
# The test compares mean completion time between the two
# instruction conditions
t_test_effect <- t.test(
  completion_time ~ condition,
  data = data_alternative,
  alternative = "two.sided"
)

# Display the results.
t_test_effect
```

```{r}
# Perform the same planned two-sided t-test on the null dataset:
# Here the true population effect is zero, so any observed
# difference between the sample means is due to random variation.
t_test_null <- t.test(
  completion_time ~ condition,
  data = data_null,
  alternative = "two.sided"
)

# Display the results.
t_test_null
```


```{r}
# Calculate the observed mean difference explicitly
# The effect is defined as:
# mean(Text + demonstration) - mean(Text-only).
effect_estimate <- mean(
  data_alternative$completion_time[data_alternative$condition == "Text + demonstration"]
) -
  mean(
    data_alternative$completion_time[data_alternative$condition == "Text-only"]
  )

# Calculate the same observed effect for the null dataset.
effect_estimate_null <- mean(
  data_null$completion_time[data_null$condition == "Text + demonstration"]
) -
  mean(
    data_null$completion_time[data_null$condition == "Text-only"]
  )

effect_estimate
effect_estimate_null
```


# Part C: Simulated Data
# 1. Introduction

```{r}
# Set the number of simulated studies
n_studies <- 10000
```

# 2. Alternative hypothesis simulation
```{r}
# Create an empty data frame to store the results from every simulated study
alternative_results <- tibble(
  study = integer(),
  p_value = numeric(),
  estimated_effect = numeric()
)

# Repeat the complete study many times
for (i in 1:n_studies) {
  
  # Generate 50 completion times for the Text-only condition
  text_times <- rnorm(
    n = n_per_group,
    mean = mean_text,
    sd = sd_completion
  )
  
  # Generate 50 completion times for the demonstration condition
  # The true mean is 30 + (-3) = 27 seconds.
  demo_times <- rnorm(
    n = n_per_group,
    mean = mean_text + effect_size,
    sd = sd_completion
  )
  
  # Prevent physically impossible negative completion times
  text_times <- pmax(text_times, 0)
  demo_times <- pmax(demo_times, 0)
  
  # Combine the two groups into one dataset
  simulated_data <- tibble(
    condition = c(
      rep("Text-only", n_per_group),
      rep("Text + demonstration", n_per_group)
    ),
    completion_time = c(
      text_times,
      demo_times
    )
  )
  
  # Run the planned two-sided Welch t-test
  test_result <- t.test(
    completion_time ~ condition,
    data = simulated_data,
    alternative = "two.sided"
  )
  
  # Calculate the estimated effect explicitly as:
  # mean(Text + demonstration) - mean(Text-only).
  estimated_effect <- mean(demo_times) - mean(text_times)
  
  # Store the p-value and estimated effect from this study
  alternative_results <- bind_rows(
    alternative_results,
    tibble(
      study = i,
      p_value = test_result$p.value,
      estimated_effect = estimated_effect
    )
  )
}
```

# 3. Null hypothesis simulation
```{r}
# Create an empty data frame to store the results from every simulated study
null_results <- tibble(
  study = integer(),
  p_value = numeric(),
  estimated_effect = numeric()
)

# Repeat the complete study many times under the null hypothesis
for (i in 1:n_studies) {
  
  # Generate 50 completion times for the Text-only condition
  text_times <- rnorm(
    n = n_per_group,
    mean = mean_text,
    sd = sd_completion
  )
  
  # Generate 50 completion times for the demonstration condition
  # Because the true effect is zero, its population mean is also 30
  demo_times <- rnorm(
    n = n_per_group,
    mean = mean_text + null_effect,
    sd = sd_completion
  )
  
  # Prevent physically impossible negative completion times.
  text_times <- pmax(text_times, 0)
  demo_times <- pmax(demo_times, 0)
  
  # Combine both conditions into one dataset.
  simulated_data <- tibble(
    condition = c(
      rep("Text-only", n_per_group),
      rep("Text + demonstration", n_per_group)
    ),
    completion_time = c(
      text_times,
      demo_times
    )
  )
  
  # Run the same two-sided Welch t-test.
  test_result <- t.test(
    completion_time ~ condition,
    data = simulated_data,
    alternative = "two.sided"
  )
  
  # Calculate the observed effect for this simulated study.
  estimated_effect <- mean(demo_times) - mean(text_times)
  
  # Store the p-value and estimated effect.
  null_results <- bind_rows(
    null_results,
    tibble(
      study = i,
      p_value = test_result$p.value,
      estimated_effect = estimated_effect
    )
  )
}
```

# 4. Type 1 Error, Type 2 Error and Power

```{r}
# Calculate the Type I error rate:
# Under the null hypothesis, a significant result (p < .05)
# represents a Type I error
type_I_error <- mean(null_results$p_value < 0.05)

type_I_error
```

```{r}
# Calculate the Type II error rate:
# This is the proportion of studies under the alternative
# hypothesis where the true effect was not detected.
type_II_error <- mean(alternative_results$p_value >= 0.05)

type_II_error
```

```{r}
# Calculate statistical power:
# Under the alternative hypothesis, a significant result
# means that the true effect was detected.
power <- mean(alternative_results$p_value < 0.05)

power
```

```{r}
# check: Type II error and power should add up to approximately 1
type_II_error + power
```

```{r}
# Create a summary table of the main simulation results:
simulation_summary <- tibble(
  measure = c(
    "Type I error",
    "Type II error",
    "Statistical power"
  ),
  estimate = c(
    type_I_error,
    type_II_error,
    power
  )
)

simulation_summary
```

# 5. P-value distributions

```{r}
# Plot the distribution of p-values under the alternative hypothesis.
# The dashed line marks the significance threshold of 0.05.
ggplot(alternative_results, aes(x = p_value)) +
  geom_histogram(bins = 20) +
  geom_vline(
    xintercept = 0.05,
    linetype = "dashed"
  ) +
  labs(
    title = "Distribution of p-values Under the Alternative Hypothesis",
    x = "p-value",
    y = "Number of simulated studies"
  ) +
  theme_minimal()
```

```{r}
# Plot the distribution of p-values under the null hypothesis
# The dashed line marks the significance threshold of 0.05:
ggplot(null_results, aes(x = p_value)) +
  geom_histogram(bins = 20) +
  geom_vline(
    xintercept = 0.05,
    linetype = "dashed"
  ) +
  labs(
    title = "Distribution of p-values Under the Null Hypothesis",
    x = "p-value",
    y = "Number of simulated studies"
  ) +
  theme_minimal()
```

# 6. Estimated-effects distributions
```{r}
# Plot the distribution of estimated effects across simulated studies
# The vertical line shows the true effect of -3 seconds:
ggplot(alternative_results, aes(x = estimated_effect)) +
  geom_histogram(bins = 30) +
  geom_vline(
    xintercept = -3,
    linetype = "dashed"
  ) +
  labs(
    title = "Distribution of Estimated Effects Across Simulated Studies",
    x = "Estimated effect (seconds)",
    y = "Number of simulated studies"
  ) +
  theme_minimal()
```

```{r}
# Plot the distribution of estimated effects under the null hypothesis
# The true effect is zero, so the distribution should be centred
# approximately around zero
ggplot(null_results, aes(x = estimated_effect)) +
  geom_histogram(bins = 30) +
  geom_vline(
    xintercept = 0,
    linetype = "dashed"
  ) +
  labs(
    title = "Distribution of Estimated Effects Under the Null Hypothesis",
    x = "Estimated effect (seconds)",
    y = "Number of simulated studies"
  ) +
  theme_minimal()
```

# 7. Different sample sizes
```{r}
sample_sizes <- c(25, 100, 200, 500)

# Number of studies used for each sample size.
n_power_studies <- 2000

# Create an empty data frame to store the results.
power_by_n <- tibble(
  n_per_group = integer(),
  power = numeric()
)

# Repeat the simulation for each sample size.
for (n_current in sample_sizes) {
  
  # Store whether each simulated study produced a significant result.
  significant_results <- numeric(n_power_studies)
  
  for (i in 1:n_power_studies) {
    
    # Generate Text-only completion times.
    text_times <- rnorm(
      n = n_current,
      mean = mean_text,
      sd = sd_completion
    )
    
    # Generate demonstration completion times under
    # the true effect of -3 seconds.
    demo_times <- rnorm(
      n = n_current,
      mean = mean_text + effect_size,
      sd = sd_completion
    )
    
    # Prevent negative completion times.
    text_times <- pmax(text_times, 0)
    demo_times <- pmax(demo_times, 0)
    
    # Combine both groups.
    simulated_data <- tibble(
      condition = c(
        rep("Text-only", n_current),
        rep("Text + demonstration", n_current)
      ),
      completion_time = c(
        text_times,
        demo_times
      )
    )
    
    # Run the two-sided Welch t-test.
    test_result <- t.test(
      completion_time ~ condition,
      data = simulated_data,
      alternative = "two.sided"
    )
    
    # Record whether the result was statistically significant.
    significant_results[i] <- test_result$p.value < 0.05
  }
  
  # Calculate power for this sample size.
  power_by_n <- bind_rows(
    power_by_n,
    tibble(
      n_per_group = n_current,
      power = mean(significant_results)
    )
  )
}

power_by_n
```

```{r}
# Visualise how statistical power changes with sample size.
ggplot(power_by_n, aes(x = n_per_group, y = power)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Statistical Power as a Function of Sample Size",
    x = "Sample size per condition",
    y = "Statistical power"
  ) +
  theme_minimal()
```