---
title: "Part B: Synthetic Data"
author: "Group 11"
date: "`r Sys.Date()`"
output: pdf_document
---

# 1. Introduction

In this part, we generate synthetic data for the proposed study.
The study investigates whether adding a human demonstration to a text
instruction affects the time an embodied AI robot needs to complete a
manipulation task.

The **independent variable** is the instruction condition:
- Text-only
- Text + demonstration

The **dependent variable** is task completion time, measured in seconds.

We generate two datasets:

1. A dataset in which a true effect exists.
2. A dataset in which the null hypothesis is true.

The effect of the human demonstration is controlled by an adjustable
parameter called `effect_size`.

# 2. Load packages

```{r setup, message=FALSE, warning=FALSE}
# Load the packages needed for data manipulation and visualisation
library(tidyverse)
```

# 3. Set seed and Define simulation parameters
```{r}
# Set a seed to for reproducibility
set.seed(42)

# Total number of robot trials
N <- 100

# Number of trials in each experimental condition
n_per_group <- N / 2

# Mean completion time for the text-only condition, in seconds
mean_text <- 30

# Standard deviation of completion time within each condition, in seconds
sd_completion <- 5

# True effect of adding a human demonstration
# A negative value means that the demonstration is expected to
# reduce completion time
# For example, -3 means that the demonstration reduces the
# population mean completion time by 3 seconds
effect_size <- -3

# Mean completion time in the demonstration condition
mean_demo <- mean_text + effect_size
```

# 4.1 Simulate the dataset with a true effect
```{r}
# Generate completion times for robots receiving the text-only instruction
# rnorm() draws random observations from a normal distribution.
# The distribution has a mean of 30 seconds and an SD of 5 seconds
text_effect <- rnorm(
  n = n_per_group,
  mean = mean_text,
  sd = sd_completion
)

# Generate completion times for robots receiving both the text instruction
# and a human demonstration
# The mean is lower by the specified effect size (-3 seconds)
demo_effect <- rnorm(
  n = n_per_group,
  mean = mean_demo,
  sd = sd_completion
)

# Combine the two groups into one data frame
data_effect <- tibble(
  condition = c(
    rep("Text-only", n_per_group),
    rep("Text + demonstration", n_per_group)
  ),
  completion_time = c(
    text_effect,
    demo_effect
  )
)

# Completion time cannot be negative in reality
# We therefore replace any negative simulated values with zero
data_effect <- data_effect %>%
  mutate(
    completion_time = pmax(completion_time, 0)
  )

# Display the first few observations
head(data_effect)
```

# 4.2 Inspect the simulated effect dataset
```{r}
# Calculate descriptive statistics for each experimental condition
data_effect %>%
  group_by(condition) %>%
  summarise(
    n = n(),
    mean = mean(completion_time),
    sd = sd(completion_time),
    min = min(completion_time),
    max = max(completion_time)
  )
```

# 5.1 Simulate the dataset under null hypothesis
```{r}
# For the null dataset, we set the true effect to zero, which means that 
# the population mean completion time is identical in the two conditions
null_effect <- 0

# Calculate the mean completion time for the demonstration condition
# under the null hypothesis
mean_demo_null <- mean_text + null_effect

# Generate completion times for the text-only condition
text_null <- rnorm(
  n = n_per_group,
  mean = mean_text,
  sd = sd_completion
)

# Generate completion times for the demonstration condition
# Since null_effect = 0, both groups have the same population mean
demo_null <- rnorm(
  n = n_per_group,
  mean = mean_demo_null,
  sd = sd_completion
)

# Combine the two groups into one data frame
data_null <- tibble(
  condition = c(
    rep("Text-only", n_per_group),
    rep("Text + demonstration", n_per_group)
  ),
  completion_time = c(
    text_null,
    demo_null
  )
)

# Ensure that completion times cannot be negative
data_null <- data_null %>%
  mutate(
    completion_time = pmax(completion_time, 0)
  )

# Display the first few observations
head(data_null)
```

# 5.2 Inspect the null dataset
```{r}
# Calculate descriptive statistics for each condition:
# Even though the population means are identical, the sample means
# will usually differ slightly because of random sampling variation
data_null %>%
  group_by(condition) %>%
  summarise(
    n = n(),
    mean = mean(completion_time),
    sd = sd(completion_time),
    min = min(completion_time),
    max = max(completion_time)
  )
```
# 6. Visualize the dataset with a true effect
```{r}
# Plot the distribution of completion times in the two conditions:

# Individual observations are shown as points
# the boxplots summarise the distribution within each condition.
ggplot(
  data_effect,
  aes(
    x = condition,
    y = completion_time
  )
) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.6) +
  labs(
    title = "Simulated Completion Time: True Effect",
    x = "Instruction condition",
    y = "Completion time (seconds)"
  ) +
  theme_minimal()
```

# 7. Visualize the null dataset
```{r}
# Visualise the completion times under the null hypothesis:
# The two groups should overlap substantially because their population
# means are identical.
ggplot(
  data_null,
  aes(
    x = condition,
    y = completion_time
  )
) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.6) +
  labs(
    title = "Simulated Completion Time: Null Effect",
    x = "Instruction condition",
    y = "Completion time (seconds)"
  ) +
  theme_minimal()
```

# 10. Compare the two conditions statistically
```{r}
# Perform the planned independent-samples two-sided t-test
# on the dataset containing a true effect:
# The test compares mean completion time between the two
# instruction conditions
t_test_effect <- t.test(
  completion_time ~ condition,
  data = data_effect,
  alternative = "two.sided"
)

# Display the results.
t_test_effect
```

```{r}
# Perform the same planned two-sided t-test on the null dataset:
# Here the true population effect is zero, so any observed
# difference between the sample means is due to random variation.
t_test_null <- t.test(
  completion_time ~ condition,
  data = data_null,
  alternative = "two.sided"
)

# Display the results.
t_test_null
```